In this chapter we will motivate the problem studied in this thesis by means of
providing historical context, as well as recent developments that are related
to the work done here.

\section{Historical Context}\label{sec:hist}
The story of Q-Balls begins in 1834 with a much more basic study of waves and
Scottish naval engineer John Scott Russell. Russell worked experimenting at the
Union Canal where he attempted to to measure the relationship between the speed
of a boat and its propelling force (which oftentimes were horses on land next
to the canal). These measurements would allow one to then convert from horse
power to steam power. One day while Russell was working he noted a rope had
caught and caused the boat to~\cite{russellsoliton}
\begin{displayquote}
suddenly stop – not so the mass of water in the channel which it had put in
motion; it accumulated round the prow of the vessel in a state of violent
agitation, then suddenly leaving it behind, rolled forward with great velocity,
assuming the form of a large solitary elevation, a rounded, smooth and well
defined heap of water, which continued its course along the channel without
change of form or diminution of speed.
\end{displayquote}
This caught Russell's attention and he followed it on horseback for nearly 2
miles before it became lost in the windings of the channel. While chasing the
wave he noted an important fact that the solitary wave preserved its shape as
it moved down the canal. So fascinated with what he had seen, Russell built a
tank to further study this phenomenon. Over the next ten years he would study
these objects and he was able to demonstrate they following qualities.
\begin{itemize}
    \item These waves were stable and could travel long
          distances\footnote{Waves you might find at a beach tend to flatten
          out over time, or peak and topple over.}
    \item The speed of the wave depends on the on the height of the wave and
          depth of the canal as \(v = \sqrt{g(d + h)}\)
    \item If a wave is too big for the depth of the water, it splits into two
          or more solitary waves
    \item Solitary waves cross each other ``without change of any kind''
\end{itemize}
As with many other great discoveries, Russell's work was not taken seriously by
the scientific community at the time because his results were not reproducible
by means of Newtonian hydrodynamics.

Nearly 60 years passed before the idea of a solitary wave became popular when
in 1895 Korteweg and de Vries~\cite{KDV} published a theory of shallow water
waves that reproduced Russell's observations and essential qualities. In
particular they found the partial differential equation
\begin{equation}
    \pdv{u}{t} + \pdv[3]{u}{x} - 6 u \pdv{u}{x} = 0
\end{equation}
which possessed solutions that fit Russell's criteria nearly exactly. This
equation possesses a solution which maintains its shape because of the simple
fact that the dispersive \(\pdv[3]{u}{x}\) term coordinates to cancel out the
effects from the nonlinear \(u\pdv{u}{x}\) term. Along with this discovery,
about 30 years later from the geometry of negatively curved
surfaces~\cite{curvedsurf}, arose the equation
\begin{equation}
    \pdv[2]{u}{x}{t} = \sin{u}
\end{equation}
and shortly after the equation
\begin{equation}
    \pdv[2]{u}{x} - \pdv[2]{u}{t} = \sin{u}
\end{equation}
coming from solid state physics~\cite{sinegordon} were both found to have
solutions that possessed solitary wave solutions. Around this point the term
``soliton'' was coined because of their particle like nature, that is their
ability to maintain their shape as they propagate, and their interaction
dynamics.

As we have seen, solitons arise from many different areas of mathematics,
physics and engineering. As we build richer and more complete models of the
universe, and parts of it, we often run into nonlinearities which have the
possibility of yielding soliton solutions should the right conditions be met.
For this reason solitons are an important object to fully understand and study.
As we will see they lead to rich areas of study for both mathematicians,
physicists, and engineers as they play a role in mathematical frameworks,
physical theories, and have many practical applications.

\section{Q-Balls}\label{sec:qballintro}
Now that we have a basic understanding of a soliton, we can can ask if there is
some sort of classification that we can do. Are all solitons essentially the
same, or do some solitons arise from fundamentally different places than others?

The answer is of course yes, and at this point we have two main categories of
solitons; \textit{topological solitons} or \textit{topological defects}, and
\textit{non-topological solitons}. As we noted in the previous section,
solitons are stable and maintain their shape. Topological solitons inherit
their stability from, not surprisingly, topological arguments. Technically,
these soliton solutions are homotopically distinct from the vacuum
state~\cite{nakahara} and hence cannot be deformed into the trivial solution.
This inability to deform them lends them a stability against decay.
Non-topological solitons however are objects which are stabilised by other
means. As is often the case, non-topological solitons are stabilised by their
``Noether Charge'' which we will learn about in Chapter~\ref{chap:fields}. This
charge stabilises them against decay and allows them to maintain their shape.

While topological defects are extremely important, and play a role in many
quantum field theories, this thesis is concerned with non-topological solitons.

The history of Q-Balls begins with the great physicist Sidney Coleman and his
paper ``Q-Balls''~\cite{coleman}, written in 1985. In this paper Coleman lays
the groundwork for the existence of Q-Balls and their stability against decay
and perturbations. Since this pioneering work there has been an enormous amount
of work done pertaining to Q-Balls and their properties, and is of course not
just limited
to~\cite{qball1,qball2,qball3,qball4,qball5,qball6,qball7,qball8,qball9}.

\section{Applications}\label{sec:apps}
Since Coleman's first paper Q-Balls have celebrated much research and activity.
Besides research into their properties, Q-Balls have been proposed as models of
new physics in many areas.

In Alexander Kusenko's work~\cite{SUSY} Q-Balls are shown to exist in
Supersymmetric generalizations of the Standard Model of Particle Physics (one
of the leading theories to go beyond the standard model). In this paper Kusenko
shows how Q-Balls could have been created at a very early time in our universe
(\(\sim1\)s). He then goes on to discuss the possible cosmological ramifications,
and in~\cite{darkmatter} Kusenko and Shaposhnikov propose Q-Balls as a candidate
for Dark Matter. In~\cite{baryogen,baryogen2}, we even see Q-Balls aiding in
the explanation of baryogenisis.\footnote{The unkown process that produced so
many more baryons (matter), than antibaryons (antimatter).}
